---
title: "Energy Data Analysis"
author: "Ashish"
date: "13/10/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Prerequisites

```{r}
# load package
library(ggplot2)
library(tidyverse)
#library(ggpubr)
#library(reshape2)
#library(MASS)
```

```{r}
#dir=getwd()
#dir
#setwd(dir)
```

## Read Data
```{r}
# read salary data and saved it in dat
dat <- read.csv("eng-data.csv", header=TRUE, sep = ",")
head(dat)
```

## Check data types
```{r}
## Convert date to Date Data Type
dat$Date <- as.Date(dat$Date)

str(dat)
```
#CHECK DIMENSION OF DATASET
```{r}
dim(dat)
```


## CHECK MISSING VALUE
```{r}
any(is.na(dat))
```
```{r}
sapply(dat, function(x) sum(is.na(x)))
```

# Remove missing values
```{r}
dat <- na.omit(dat)
any(is.na(dat))
```
## Summary statistics
```{r}
summary(dat)

```

## Plot Date Vs (Consumption, Wind and Solar)

```{r}
library(ggpubr)
p1 <- ggplot(dat, aes(x =Date, y =Consumption)) + geom_line(alpha=0.5, size=0.5, color="red") + labs(title="Date vs Consumption")
p2 <- ggplot(dat, aes(x =Date, y =Wind)) + geom_line(alpha=0.5, size=0.3, color="blue") + labs(title="Date vs Wind")
p3 <- ggplot(dat, aes(x =Date, y =Solar)) + geom_line(alpha=0.5, size=0.3) + labs(title="Date vs Solar")
ggarrange(p1, p2, p3, ncol = 2, nrow = 2)

```
## Lets investigate a single year to have a closer look.
## Add a new Date column into an existing data frame and split it
```{r}
# apply separate function to split x into Year", "Month", "Day
x <- dat$Date
dat$date_x <- as.Date(x)

# apply separate function to split x into Year", "Month", "Day
dat_2 <- separate(dat, date_x, c("Year", "Month", "Day"), sep = '-')
head(dat_2)
```
## Change date items (year, month, date) seperated into numeric
```{r}
## Save data for group by Analysis
dat_3 <- dat_2

dat_2$Year  <- as.numeric(dat_2$Year)
dat_2$Month <- as.numeric(dat_2$Month)
dat_2$Day   <- as.numeric(dat_2$Day)
head(dat_2)
```

## Plot Consumption vs a Year, say 2016
```{r}
# filter data for Year 2016
dy <- filter(dat_2, dat_2$Year == 2016)
ggplot(dy, aes(x=Date, y=Consumption)) + geom_line(alpha=0.5, size=0.5, color="red") + labs(title="Plot Consumption vs a Year, say 2016")
```

####From the above figure we can clearly see the consumption of electricity for Year 2016. There is a drastic decrease in the consumption of electricity at the end of the year (December) and during August.

#Lets investigate December month in 2016
##Plot Consumption for December month 2016

```{r}
# filter data for month December and Year 2016
dy <- filter(dat_2, dat_2$Year == 2016 & Month == 12)
ggplot(dy, aes(x=Date , y=Consumption)) + geom_line(alpha=0.5, size=0.5, color="red") + geom_point()+ labs(title="Plot Consumption for December month 2016")

```
####As shown in the above figure, electricity consumption is higher on weekdays and lowest at the weekends. We can also investigate the consumption for each day of the month.

#Lets investigate consumption in the last week of December month in 2016
##Plot Consumption for December month 2016 between date 23 and 26 (Christmas Time)

```{r}
# filter data for last week of December 2016
dy <- filter(dat_2, dat_2$Year == 2016 & Month == 12 & Day >=23 & Day <=30)
ggplot(dy, aes(x=Date , y=Consumption)) + geom_line(alpha=0.5, size=0.5, color="red") + geom_point()+ labs(title="Plot Consumption for December month 2016 between date 23 and 26 (Christmas Time)")
```

#### From the above fiure, we can see that electricity consumption was lowest on the day of Christmas, probably because people were busy partying. After Christmas, the consumption increased.

## Grouping time series data We can group the data by different time periods and present them in box plot.
```{r}
### Plot data by months. Use the box plots to visualise the data vs month
pa1 <- ggplot(data=dat_3, mapping=aes(x=Month, y=Consumption, fill=Month)) + geom_boxplot(alpha=0.9) + theme(legend.position="none") + labs(title="Date vs Consumption By Month")

pa2 <- ggplot(data=dat_3, mapping=aes(x=Month, y=Solar, fill=Month)) + geom_boxplot(alpha=0.9)+ theme(legend.position="none") + labs(title="Date vs Solar By Month")

pa3 <- ggplot(data=dat_3, mapping=aes(x=Month, y=Wind, fill=Month)) + geom_boxplot(alpha=0.9)+ theme(legend.position="none") + labs(title="Date vs Wind By Month")

ggarrange(pa1, pa2, pa3 , ncol = 1, nrow = 3) #labels = c("Consumption", "Wind", "Solar")
```
#Lets investigate consumption of electricity by the day of the week

##Convert Date into Weekday using weekdays Function

```{r}
# add new column and convert Date into weekdays
dat_2$Weekday <- weekdays(dat_2$Date)

### Plot data by months. Use the box plots to visualise the data vs month
pa1 <- ggplot(data=dat_2, mapping=aes(x=Weekday, y=Consumption, fill=Weekday)) + geom_boxplot(alpha=0.9) + theme(legend.position="none") + labs(title="Date vs Consumption By Weekday")

pa2 <- ggplot(data=dat_2, mapping=aes(x=Weekday, y=Solar, fill=Weekday)) + geom_boxplot(alpha=0.9)+ theme(legend.position="none") + labs(title="Date vs Solar By Weekday")

pa3 <- ggplot(data=dat_2, mapping=aes(x=Weekday, y=Wind, fill=Weekday)) + geom_boxplot(alpha=0.9) + theme(legend.position="none") + labs(title="Date vs Wind By Weekday")

ggarrange(pa1, pa2, pa3 , ncol = 1, nrow = 3) #labels = c("Consumption", "Wind", "Solar")
```
#### The above figure shows that electricity consumption is higher on weekdays than on weekends. Interestingly, there are more outliers on the weekdays.

## Plot Consumption vs Wind.Solar of 2016
```{r}
# filter data for last week of December 2016
dy2 <- filter(dat_2, Year == 2016)
ggplot(dy2, aes(x=Date)) + geom_line(aes(y = Wind.Solar), color = "darkred", size=0.8) +
geom_line(aes(y = Consumption), color="steelblue", linetype="twodash")


```


