---
title: "MT Cars Analysis"
author: "Ashish"
date: "14/10/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Prerequisites: install packages

```{r}
# load package
library(ggpubr)
library(PerformanceAnalytics)
library(corrplot)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(datasets)
library(dplyr)

```

```{r}
#dir=getwd()
#dir
#setwd(dir)
```

##R functions for Correlation Correlation coefficient can be computed
```{r}
#using the functions cor() or cor.test().
#cor(x, y, method = c("pearson", "kendall", "spearman"))
#cor.test(x, y, method=c("pearson", "kendall", "spearman"))
```


## Import data into R
```{r}
# read salary data and saved it in dat
dat <- mtcars
head(dat, 10)
```

## Inspect the data: check data format and missing values

```{r}
str(dat)
```

#CHECK DIMENSION OF DATASET
```{r}
dim(dat)
```


## CHECK MISSING VALUE
```{r}
any(is.na(dat))
```

## Summary statistics
```{r}
summary(dat)
```

## Identify which variables do you want to check the Correlation between them
#### correlation between mpg and wt variables


### Visualise the variables using scatter plots, ggpubr R package
```{r}
ggscatter(dat, x = "mpg", y = "wt", add = "reg.line", conf.int = TRUE, xlab = "Miles/gallon", ylab = "Weight")
#ggplot(dat, aes(x=mpg , y=wt)) + geom_point(color="red") + geom_smooth(method = lm)
```

## Pearson correlation, Spearman correlation and Kendall correlation
```{r}
ggscatter(dat, x="mpg", y="wt", add="reg.line", conf.int=TRUE, cor.coef=TRUE, cor.method="pearson", xlab="Miles/gallon", ylab="Weight")
```
#### From the plot, we can see that based on the Covariation: the relationship between mpg and wt is linear.

## Normality test - Are mpg and wt follow a normal distribution?


```{r}
# The Prob < W value listed in the output is the p-value. If the chosen alpha level is 0.05 and the p-value is less than 0.05, then the null hypothesis that the data are normally distributed is rejected. If the p-value is greater than 0.05, then the null hypothesis is not rejected.

# Shapiro test for mpg
shapiro.test(dat$mpg)

# Shapiro test for wt
shapiro.test(dat$wt)
```
##### From the results, we can see that p-values are greater than the significance level 0.05. This means H1 will be rejected. Hence, we can assume mpg and wt are normally distributed.


## Pearson correlation
```{r}
#DEFAULT HYPOTHESIS IS RELATION DOESN'T EXIST. p-value > 0.05 correlation doesn't exists (h1 hypothsis), p-value < 0.05 correlation exists
res <- cor.test(dat$wt, dat$mpg, method = "pearson")
res
```
## Spearman rank correlation coefficient
```{r}

res2 <- cor.test(dat$wt, dat$mpg, method = "spearman")
res2
```
## Kendall rank correlation test
```{r}
res3 <- cor.test(dat$wt, dat$mpg, method="kendall")
res3
```
## correlation between the possible pairs of variables
#### correlation between the possible pairs of variables

```{r}
res4 <- round(cor(dat), 2)
res4
```
## Visualisation of the correlation matrix.
#### We can use chart.Correlation to visualise the correlation matrix

```{r}
chart.Correlation(dat, histogram=TRUE, pch="19")
```
## Visualisation of the correlation matrix - Circle, Pie, Number
```{r}
#We can use corrplot function. corrplot is a graph of correlation matrix. In this plot, correlation coefficients is colored according to the value.

M <- cor(dat)
corrplot(M, method = "circle")
corrplot(M, method = "circle", type = "upper")
corrplot(M, method="pie")
corrplot(M, method="number")
```
Correlation Applications
• Prediction: If there is a relationship between two variables, we can make predictions about one from another.
• Validity: correlation validity between a new measure and an established measure.
• Reliability: measures consistent or observers consistent.
• Theory verification: Predictive validity

```{r}

```






